High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2021-05-27 Draft_V4046Sgr_mnras.tex
Updates from Overleaf

The simulated 1.3\,mm continuum image reproduces the two observed rings: the faint Ring13 and the brighter Ring24. The introduced Ring5 is not visible in the simulated image as its predicted peak intensity is around two times the noise in the ALMA image ($\sim$1$\times10^{-7}$\,Jy\,$\mathrm{beam}^{-1}$). This gives us an upper limit for the total millimetre-sized dust mass present in Ring5 of $\sim$2$\times10^{-5}$\,M$_{\earth}$. The depletion of large dust grains in Ring5 is consistent with efficient dust trapping in Ring13. A zone of radially increasing gas pressure can entirely filter out large dust grains from inner regions, while smaller grains can be able to overcome this barrier due to their strong frictional coupling to the gas accretion flow \citep[studied in the context of planetary gaps,][]{Rice2006,Zhu2012,Weber2018}. Still, the strong depletion of large particles demands that within Ring5 dust evolution is extremely inefficient or limited, or that large particles are rapidly removed. Otherwise, the present small grains should quickly coagulate to form detectable grain sizes in the ALMA observation in Fig.~\ref{fig:images_vs_simulated} \citep{Drazkowska2019}. The central emission is not present in the simulated 1.3 mm continuum image as it was impossible to achieve a bright blob without disrupting the SED fit (read below).

As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig.~\ref{fig:radprofiles}), we can assume that the model provides a possible approximation of the disc structure, including the dimensions of Ring13. Accordingly, taking the parametric model values for Ring13, it has a radius of 13.22\,au, a FWHM of 2.00\,au, and a scale height FWHM of the millimetre-sized dust of 0.6\,au. The total dust mass of Ring13 is about 0.7\,M$_{\earth}$ in the model. In turn, our model reproduces the observations of Ring24 with peak intensity at $\sim$30\,au, and break at $\sim$36\,au. It contains a total dust mass of $\sim$45\,M$_{\earth}$, that stands under the lower limit of 60\,M$_{\earth}$ inferred by \citet{Ru_z_Rodr_guez_2019} from a previous 860\,$\micron$ continuum ALMA image. The model predictions for the millimetre-sized dust population in Ring13 are close to the measurements, with only a 0.07\,au difference in the centroid location of the ring, and a 0.46\,au difference in the width estimation. Since the scale height FWHM of the large grains in Ring13 is 0.6\,au, and has a measured width of 2.46$\pm$0.40\,au (See sec.~\ref{subsec:ALMA}), the ring is 4.1$\pm$0.9 times more extended radially than vertically. The central emission radial profile is also well reproduced, and the lack of a high NIR excess indicate that the dust in this blob must be at least bigger than 0.5\,mm (see Appendix~\ref{sec:Appendix}). In addition, our interpretation of a Gaussian close ring agrees with the mass estimates of \citet{Francis_2020} with a dust mass of 0.012\,M$_{\earth}$, implying a slightly higher lower limit for the dust population around the central region.

@@

The effective beam size of the presented ALMA observations marginally resolves the FWHM of Ring13 to be $w_\mathrm{d} = 2.46\pm0.40$\,au. Using that the FWHM of the gas in Ring13 in the RT model corresponds to $w_\mathrm{g} = 5.42^{+0.01}_{-1.15}$\,au, we can see that large dust grains have to be radially converged to explain the discrepancy.

@@

The model presented here has many degrees of freedom and each model requires a high level of computational power and time, so a MCMC optimization or similar methods of parameter exploration are impossible to acquire. But as a way to compute a measure of goodness of fit and to try to quantify the improvements we did a partial exploration of the parameter space. We explore $\chi^2$ space in the vicinity of the model values for the scale height at $r_\circ$, the location of Ring5, the width of the gas component of Ring13, the width of the gas part of Ring5, the width of the large dust portion of Ring13 and around the lower limit of the grain sizes in the blob. The parameter space exploration is shown in Figure~\ref{fig:chi}. Our model fits the SED and two images, so the total $\chi^2$ value for a given model is composed of the sum of the $\chi^2$ values for each of the three fits.

@@

For the the other three parameters, we found that although our chosen values are close to a minimum, the geometries are more suggestive of different border conditions. These exploration allow us to support the idea of Ring5, as it is clear that the absence of it creates a great increment in the total $\chi^2$ value, but also that the ring must be quite thin. For the case of the lower limit of the central blob, it is evident that there is a threshold around 500\,$\micron$, where the SED start to deviate far away from a minimum, that point represent where the NIR excess becomes significant.

>>>>

The simulated 1.3\,mm continuum image reproduces the two observed rings: the faint Ring13 and the brighter Ring24. The introduced Ring5 is not visible in the simulated image as its predicted peak intensity is around two times the noise in the ALMA image ($\sim$1$\times10^{-7}$\,Jy\,$\mathrm{beam}^{-1}$). This gives us an upper limit for the total millimetre--sized dust mass present in Ring5 of $\sim$2$\times10^{-5}$\,M$_{\earth}$. The depletion of large dust grains in Ring5 is consistent with efficient dust trapping in Ring13. A zone of radially increasing gas pressure can entirely filter out large dust grains from inner regions, while smaller grains can be able to overcome this barrier due to their strong frictional coupling to the gas accretion flow \citep[studied in the context of planetary gaps,][]{Rice2006,Zhu2012,Weber2018}. Still, the strong depletion of large particles demands that within Ring5 dust evolution is extremely inefficient or limited, or that large particles are rapidly removed. Otherwise, the present small grains should quickly coagulate to form detectable grain sizes in the ALMA observation in Fig.~\ref{fig:images_vs_simulated} \citep{Drazkowska2019}. The central emission is also well reproduced with the inclusion of a Gaussian close ring, resulting in the observed central blob. The lack of a high NIR excess in the SED indicate that the dust in this blob must be at least bigger than 0.5\,mm (see Appendix~\ref{sec:Appendix}), but we choose to use 0.8\,mm as the lower limit for the dust population that composes this feature because this predicts a dust mass of 0.012\,M$_{\earth}$, that is in agreement with the mass estimate of \citet{Francis_2020} (0.013$\pm$0.002\,M$_{\earth}$).

As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig.~\ref{fig:radprofiles}), we can assume that the model provides a possible approximation of the disc structure, including the dimensions of Ring13. The FWHM of the gas in Ring13 in the RT model corresponds to $w_\mathrm{g} = 5.18^{+0.32}_{-0.20}$\,au, but the  large dust grains have to be radially converged to explain the ALMA image. Accordingly, taking the parametric model values for the large dust in Ring13, it has a FWHM of 2.00\,au, a radius of 13.22\,au, and a scale height FWHM of the millimetre-sized dust of 0.25\,au. The total dust mass of Ring13 would be about 0.7\,M$_{\earth}$. The model predictions for the millimetre-sized dust population in Ring13 are close to the measurements, with only a 0.07\,au difference in the centroid location of the ring, and a 0.46\,au difference in the width estimation. Since the scale height FWHM of the large grains in Ring13 is 0.25\,au, and that from the ALMA observation we can resolve a width of 2.46$\pm$0.40\,au (See sec.~\ref{subsec:ALMA}), the ring is 9.7$\pm$1.5 times more extended radially than vertically. Looking at the rest of the disc, in turn, our model reproduces the observations of Ring24 with peak intensity at $\sim$30\,au, and breaks at $\sim$36\,au. It contains a total dust mass of $\sim$45\,M$_{\earth}$.%, that stands under the lower limit of 60\,M$_{\earth}$ inferred by \citet{Ru_z_Rodr_guez_2019} from a previous 860\,$\micron$ continuum ALMA image.

@@

@@

The model presented here has many degrees of freedom and each model requires a high level of computational power and time, so a MCMC optimization or similar methods of parameter exploration are impossible to acquire. But as a way to compute a measure of goodness of fit and to try to quantify the improvements, we did a partial exploration of the parameter space. We explore $\chi^2$ space in the vicinity of the model values for the scale height at $r_\circ$, the location of Ring5, the width of the gas component of Ring13, the width of the gas part of Ring5, the width of the large dust portion of Ring13 and around the lower limit of the grain sizes in the blob. The parameter space exploration is shown in Figure~\ref{fig:chi}. Our model fits the SED and two images, so the total $\chi^2$ value for a given model is composed of the sum of the $\chi^2$ values for each of the three fits.
\begin{equation}
    \chi^2_{\mathrm{total}}=\chi^2_{\mathrm{SED}}+\chi^2_{\mathrm{ALMA}}+\chi^2_{\mathrm{SPHERE}}
\end{equation}

@@

For the the other three parameters, we found that although our chosen values are close to a minimum, the geometries are more suggestive of different border conditions. These exploration allow us to support the idea of Ring5, as it is clear that the absence of it creates a great increment in the total $\chi^2$ value, but also that the ring must be quite thin. For the case of the lower limit of the central blob, it is evident that there is a threshold around 500\,$\micron$ where the SED start to deviate far away from the typical value. That point represent where the NIR excess becomes significant.
